# Activity - Course Project 1

```{r, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, fig.path="figure/")
```

## 1 - Loading and Preprocessing the Data

```{r, echo = TRUE}
library(ggplot2)
library(dplyr)

setwd("C:/Users/danie/Desktop/Programming/John Hopkins Data Specialisation Course/Module 5/Course Project 1")
activity <- read.csv("activity.csv")

activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
```

## 2 - What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day.
2. Make a histogram of the total number of steps taken each day.
3. Calculate and report the mean and median of the total number of steps taken per day.

```{r}
steps_per_day <- aggregate(steps ~ date, activity, FUN = sum, na.rm = TRUE)

ggplot(steps_per_day, aes(x=steps)) + geom_histogram(fill='blue', color='black', bins=10)

paste('Mean is:', mean(steps_per_day$steps))
paste('Median is:', median(steps_per_day$steps))
```

## 3 - What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps_per_interval <- aggregate(steps ~ interval, activity, FUN = mean, na.rm=TRUE)
ggplot(steps_per_interval, aes(x = interval, y = steps)) + geom_line()

paste('Interval with the max average steps is:',steps_per_interval[which.max(steps_per_interval$steps),]$interval)
```

## 4 - Imputing missing values

1. Calculate and report the total number of missing values in the dataset.
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. 
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
paste('The total number of rows with NA values is:',sum(is.na(activity$steps)))

activityCopy <- activity

for(i in 1:nrow(activityCopy)){
    if(is.na(activityCopy$steps[i])){
        
        theInterval = activityCopy$interval[i]
        
        activityCopy$steps[i] <- steps_per_interval[steps_per_interval$interval == theInterval,]$steps
    }
}

steps_per_day_copy <- aggregate(steps ~ date, activityCopy, FUN=sum)
ggplot(steps_per_day_copy, aes(x=steps)) + geom_histogram(fill='blue', color='black', bins=10)

paste('Mean is: ', mean(steps_per_day_copy$steps))
paste('Median is: ', median(steps_per_day_copy$steps))
```

The mean value has not changed at all when filling in missing values using the mean for the relevant 5 minute interval. The median has changed by the smallest amount compared to before (1 step).

## 5 - Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
activityCopy$day <- weekdays(activityCopy$date)

for(i in 1:nrow(activityCopy)){
    if(activityCopy$day[i] %in% c("Saturday", "Sunday")){
        activityCopy$day[i] <- "Weekend"
    }
    else{
        activityCopy$day[i] <- "Weekday"
    }
}

steps_per_day <- aggregate(activityCopy$steps ~ activityCopy$interval + activityCopy$day, activityCopy, mean)
colnames(steps_per_day) <- c("interval", "day", "steps")

ggplot(steps_per_day, aes(x = interval, y = steps)) + geom_line() + facet_grid(day~.)
```